Stellar multiplicity in \Gaia

### 2018-05-20 article/ms.tex
words

\documentclass[twocolumn]{aastex61}
\usepackage{bm}

@@

\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\acronym}[1]{{\small{#1}}}
\newcommand{\Gaia}{\project{Gaia}}
\newcommand{\gaia}{\project{gaia}}

@@

@@

\title{Stellar multiplicity}

@@

@@

\begin{abstract}
Velociraptors can't see you if you don't move. (\Gaia can see you if you can.)
\end{abstract}

@@

\section{Model}

@@

\software{numpy, scipy, astropy, Stan}

\acknowledgements
- If these peeps are not co-authors: Hogg, Spina, Schlaufman.
- Gaia/DPAC
- Gaia archive?
- Gaia DR2 hack week at Flatiron
- ARC DP

%This work has made use of CosmoHub.

%CosmoHub has been developed by the Port d'Informació Científica (PIC), maintained through a collaboration of the Institut de Física d'Altes Energies (IFAE) and the Centro de Investigaciones Energéticas, Medioambientales y Tecnológicas (CIEMAT), and was partially funded by the "Plan Estatal de Investigación Científica y Técnica y de Innovación" program of the Spanish government.

>>>>

\documentclass[twocolumn]{aastex61}
\usepackage{bm}
\usepackage{color}

@@

\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\package}[1]{\texttt{#1}}
\newcommand{\acronym}[1]{{\small{#1}}}
\newcommand{\Gaia}{\project{Gaia}}
\newcommand{\gaia}{\project{gaia}}
\newcommand{\todo}[1]{\textcolor{red}{#1}}

\newcommand{\GaiaRVE}{\sigma_{\mathrm{V}_\mathrm{R}}^\mathrm{MTA}}
\newcommand{\RVJitter}{\sigma(\mathrm{V}_\mathrm{R}^{t})}

@@

%\graphicspath{../figures}

@@

\title{Stellar multiplicity in \Gaia}

@@

%\author{Daniel Foreman-Mackey}
%\affiliation{Flatiron Institute, 162 Fifth Ave, New York, NY 10010, USA}

@@

\begin{abstract}
Most stars are in binary systems. 
The occurrence rate of binary star systems is critically important to many 
fields of astrophysics, but the effects of binaries are commonly ignored in 
astronomical data analysis applications.
Here we construct a model to probabilistically identify binary stars based on
excess radial velocity scatter, photometry, and astrometric excess noise.

We identify binaries \todo{within some orbital and SP limits} and show that 
the orbital semi-amplitude is well-approximated by the excess radial velocity
scatter for systems within these criteria.

\todo{We make a few interesting tests about binarity...}
\end{abstract}

@@

\section{Method}

There are at least three ways that binary systems can be reliably
identified using from \Gaia data: through radial velocity variations, 
anomalous photometric colours, or excess astrometric noise. In this 
work we do not make use of astrometric noise because of the 
degree-of-freedom astrometry bug present in the second \Gaia\ data 
release \citep{Lindegren:2018}. We will describe the radial velocity
information that forms part of our model for identifying binary stars,
and then discuss the photometric information used, before explicitly
outlining the assumptions of our model.

The second \Gaia\ data release provides median radial velocity 
estimates for 7,224,631 stars. The radial velocity error provided
($\GaiaRVE$; \texttt{radial\_velocity\_error}) is a function of the 
number of transits $N$ (i.e., observations; column 
\texttt{rv\_nb\_transits}) and the scatter among those $N$ measurements. 
This allows us to calculate the scatter among those measurements, or
rather the \emph{radial velocity jitter}, for each source
\begin{eqnarray}
\RVJitter = \sqrt{\frac{2N}{\pi}}\GaiaRVE \quad .
\end{eqnarray}

For a single star system observed $N$ times, the source radial velocity
jitter represents either the minimum uncertainty with which
\Gaia\ can measure radial velocity for a star of that apparent 
brightness and colour. However for stars with stellar companions,
the radial velocity scatter is a function of orbital parameters 
and observing epochs. Thus we can reasonably assume that the
radial velocity jitter will be a mixture of
two populations: single stars, where the radial velocity jitter
represents the minimum radial velocity noise that \Gaia\ can
measure for a star of that type, and a binary population that
could have radial velocity variances well above the noise floor.

\begin{figure*}
	\includegraphics[width=1.0\textwidth]{../figures/sb9-comparison.pdf}
    \caption{Source radial velocity variance shown with respect to flux in
    the \Gaia\ BP, G, and RP bands. The source radial velocity variance 
    most closely follows the RP flux as it is closer to the \Gaia\ RVS
    band.}
\end{figure*}

\begin{figure}
	\includegraphics[width=0.5\textwidth]{../figures/sb9-comparison.pdf}
    \caption{Source mean \Gaia\ RP flux and radial velocity variance.
    Points are represented by their mean $BP-RP$ color, confirming that
    there is some color dependency on the source radial velocity variance.}
    \label{fig:rp-flux-color}
\end{figure}

\begin{figure}
	\includegraphics[width=0.5\textwidth]{../figures/sb9-comparison.pdf}
    \caption{Mixture model fitted.}
    \label{fig:mixture-model}
\end{figure}

\begin{figure}
	\includegraphics[width=0.5\textwidth]{../figures/sb9-comparison.pdf}
    \caption{Detected binary fraction as a function of the mean RP flux
    (top panels), the $BP-RP$ color (second panels), the number of 
    radial velocity transits (third panels), and the inferred temperature
    (bottom panels). No trend is observed with mean RP flux or the number 
    of radial velocity transits. However, our model finds bluer (hotter)
    stars are more likely to be binaries, despite including color as a
    dependency in our model.}
    \label{fig:binary-fraction-wrt-params}
\end{figure}

% we assume what?
%

% given these assumptions, the model we adopt is XYZ
% we chose to model in RP flux as it more closely represents the flux counts that would be received by the Gaia RVS, rather than G or BP flux.

% the model we use is

%

@@

\acknowledgements

It is a pleasure to thank
	David W. Hogg,
	Lorenzo Spina, 
		and 
	Kevin C. Schlaufman.
This work was partly supported through the Australian Research Council 
through Discovery Grant DP160100637.
This work has made use of data from the European Space Agency (ESA) mission {\it
Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed by the {\it Gaia} Data
Processing and Analysis Consortium (DPAC,
\url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding for the DPAC
has been provided by national institutions, in particular the institutions
participating in the {\it Gaia} Multilateral Agreement.  This research was
started at the NYC Gaia DR2 Workshop at the Center for Computational
Astrophysics of the Flatiron Institute in 2018 April.

This work has made use of CosmoHub. CosmoHub has been developed by the Port d'Informaci\'o Cient\'ifica (PIC), maintained through a collaboration of the Institut de F\'isica d'Altes Energies (IFAE) and the Centro de Investigaciones Energ\'eticas, Medioambientales y Tecnol\'ogicas (CIEMAT), and was partially funded by the "Plan Estatal de Investigaci\'on Científica y T\'ecnica y de Innovaci\'on" program of the Spanish government.

\software{
	\package{Astropy} \citep{astropy},
    \package{IPython} \citep{ipython},
    \package{matplotlib} \citep{mpl},
    \package{numpy} \citep{numpy},
    \package{scipy} \citep{scipy},
    \package{Stan} \citep{stan},
    \package{CosmoHub} \citep{cosmohub}
    %tensorflow
}

\bibliographystyle{aasjournal}
\bibliography{velociraptor}
